Testing the Evolution of the Correlations between Supermassive Black Holes and their Host Galaxies using Eight Strongly Lensed Quasars

### 2020-09-25 manuscript/manuscript.tex
The final submission

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[X. Ding et al.]{
Xuheng Ding,$^{1}$\thanks{E-mail: dxh@astro.ucla.edu}
Tommaso Treu,$^{1}$
Simon Birrer,$^{2}$
Adriano Agnello,$^{3}$\newauthor
Dominique Sluse,$^{4}$
Chris Fassnacht,$^{5}$
Matthew W. Auger,$^{6}$
Kenneth C. Wong,$^{7}$ \newauthor
Sherry H. Suyu,$^{8,9,10}$
Takahiro Morishita,$^{11}$
Cristian E. Rusu,$^{12,13,14}$
Aymeric Galan,$^{15}$
\\
\\
% List of institutions
$^{1}$Department of Physics and Astronomy, University of California, Los Angeles, CA, 90095-1547, USA\\
$^{2}$Kavli Institute for Particle Astrophysics and Cosmology and Department of Physics, Stanford University, Stanford, CA 94305, USA\\
$^{3}$European Southern Observatory, Karl-Schwarzschild-Strasse 2, 85748 Garching, Germany\\
$^{4}$STAR Institute, Quartier Agora - All\'ee du six Ao\^ut, 19c B-4000 Li\`ege, Belgium\\
$^{5}$Department of Physics, University of California, Davis, CA 95616, USA\\
$^{6}$Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK\\
$^{7}$Kavli IPMU (WPI), UTIAS, The University of Tokyo, Kashiwa, Chiba 277-8583, Japan\\
$^{8}$Max-Planck-Institut f{\"u}r Astrophysik, Karl-Schwarzschild-Str.~1, 85748 Garching, Germany\\
$^{9}$Physik-Department, Technische Universit\"at M\"unchen, James-Franck-Stra\ss{}e~1, 85748 Garching, Germany\\
$^{10}$Academia Sinica Institute of Astronomy and Astrophysics (ASIAA), 11F of ASMAB, No.1, Section 4, Roosevelt Road, Taipei 10617, Taiwan\\
$^{11}$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA\\
$^{12}$National Astronomical Observatory of Japan, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan\\
$^{13}$Subaru Telescope, National Astronomical Observatory of Japan, 650 N Aohoku Pl, Hilo, HI 96720\\
$^{14}$Department of Physics, University of California, Davis, 1 Shields Avenue, Davis, CA 95616, USA\\
$^{15}$Institute of Physics, Laboratory of Astrophysics, Ecole Polytechnique 
F\'ed\'erale de Lausanne (EPFL), Observatoire de Sauverny, 1290 Versoix, 
Switzerland
}

@@

\subsection{The \mbh-\mstar\ relation}\label{sec:relation}
In Figure~\ref{fig:scaling_relation}-(left), we plot \mbh\ vs \mstar\ for our sample, together with the comparison sample introduced in Section~\ref{sec:sample_select}.  We expect that the uncertainty of the \mbh\ (i.e., $0.4~$dex) dominates the error budget for the entire sample. Following D20, we adopt, as a baseline, a local \mbh-\mstar\ of the form:
\begin{eqnarray}
\label{eq:MMlocal}
\log \big( \frac{\mathcal M_{\rm BH}}{10^{7}M_{\odot}})= \alpha + \beta \log(\frac{M_*}{10^{10}M_{\odot}}),
  \end {eqnarray}
with  $\alpha = 0.27$, $\beta = 0.98$ based on the local sample of 55 objects measured in a consistent manner. We find that our lensed systems are above the local relation, consistent with the inference from the 32  non-lensed AGNs published by D20 in a  similar redshift range. To quantify the evolution as a function of redshift, we parameterize it as:
\begin{eqnarray}
\label{eq:offset}
\Delta \log \mathcal M_{\rm BH}= \gamma \log (1 + z),
\end{eqnarray} 
where $\Delta \log \mathcal M_{\rm BH}$ is the offset of \mbh\ with respect to the local baseline at fixed \mstar. To make a direct comparison, we reproduce the plot shown in Figure~8 of D20. Then, we add our new lensed AGN measurements and show the result in Figure~\ref{fig:scaling_relation}-(right). We find that the offset from the local relationship is similar for the two samples. Based on Equation~\eqref{eq:offset}, we fit the evolution of the eight lensed systems and obtain $\gamma=1.05\pm0.44$, which agrees with the results of D20 ($\gamma=1.03\pm0.25$ using 32 AGN) within $1-\sigma$ level. The consistency between the two measurements provides an important verification of the accuracy of the results and strengthens the conclusions drawn by D20 that the {\it observed} value of \mbh\ at a fixed \mstar\ tends to be larger at higher redshift than the ones in the local universe. {\bf It is possible that the uncertainty of the \mstar\ could be higher than our estimation (i.e., 0.2~dex), given that the star formation histories could vary widely at high redshift ($z\sim$2). However, even when our stellar mass uncertainty is boosted, the evolution inferred using the eight lensed systems is still robust --- $\gamma=1.05\pm0.48$ assuming  $\Delta$\mstar = 0.3~dex, and $\gamma=1.03\pm0.53$ assuming $\Delta$\mstar = 0.4~dex.}

@@

This work is based in part on observations made with the NASA/ESA Hubble Space Telescope, obtained at the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5-26555. X.D., S.B., and T.T. acknowledge support by the Packard Foundation through a Packard Research fellowship to T.T. This work is supported by JSPS KAKENHI Grant Number JP18H01251 and the World Premier International Research Center Initiative (WPI), MEXT, Japan. This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No 787886).
C.D.F. acknowledge support for this work from the National Science Foundation under Grant Numbers AST-1312329 and AST-1907396.
SHS thanks the Max Planck Society for support through the Max Planck Research Group.

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[X. Ding et al.]{
Xuheng Ding,$^{1}$\thanks{E-mail: dxh@astro.ucla.edu}
Tommaso Treu,$^{1}$
Simon Birrer,$^{2}$
Adriano Agnello,$^{3}$\newauthor
Dominique Sluse,$^{4}$
Chris Fassnacht,$^{5}$
Matthew W. Auger,$^{6}$
Kenneth C. Wong,$^{7}$ \newauthor
Sherry H. Suyu,$^{8,9,10}$
Takahiro Morishita,$^{11}$
Cristian E. Rusu,$^{12,13,14}$
Aymeric Galan,$^{15}$
\\
\\
% List of institutions
$^{1}$Department of Physics and Astronomy, University of California, Los Angeles, CA, 90095-1547, USA\\
$^{2}$Kavli Institute for Particle Astrophysics and Cosmology and Department of Physics, Stanford University, Stanford, CA 94305, USA\\
$^{3}$European Southern Observatory, Karl-Schwarzschild-Strasse 2, 85748 Garching, Germany\\
$^{4}$STAR Institute, Quartier Agora - All\'ee du six Ao\^ut, 19c B-4000 Li\`ege, Belgium\\
$^{5}$Department of Physics {\bf and Astronomy}, University of California, Davis, CA 95616, USA\\
$^{6}$Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK\\
$^{7}$Kavli IPMU (WPI), UTIAS, The University of Tokyo, Kashiwa, Chiba 277-8583, Japan\\
$^{8}$Max-Planck-Institut f{\"u}r Astrophysik, Karl-Schwarzschild-Str.~1, 85748 Garching, Germany\\
$^{9}$Physik-Department, Technische Universit\"at M\"unchen, James-Franck-Stra\ss{}e~1, 85748 Garching, Germany\\
$^{10}$Academia Sinica Institute of Astronomy and Astrophysics (ASIAA), 11F of ASMAB, No.1, Section 4, Roosevelt Road, Taipei 10617, Taiwan\\
$^{11}$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA\\
$^{12}$National Astronomical Observatory of Japan, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan\\
$^{13}$Subaru Telescope, National Astronomical Observatory of Japan, 650 N Aohoku Pl, Hilo, HI 96720\\
$^{14}$Department of Physics, University of California, Davis, 1 Shields Avenue, Davis, CA 95616, USA\\
$^{15}$Institute of Physics, Laboratory of Astrophysics, Ecole Polytechnique 
F\'ed\'erale de Lausanne (EPFL), Observatoire de Sauverny, 1290 Versoix, 
Switzerland
}

@@

\subsection{The \mbh-\mstar\ relation}\label{sec:relation}
In Figure~\ref{fig:scaling_relation}-(left), we plot \mbh\ vs \mstar\ for our sample, together with the comparison sample introduced in Section~\ref{sec:sample_select}.  We expect that the uncertainty of the \mbh\ (i.e., $0.4~$dex) dominates the error budget for the entire sample. Following D20, we adopt, as a baseline, a local \mbh-\mstar\ of the form:
\begin{eqnarray}
\label{eq:MMlocal}
\log \big( \frac{\mathcal M_{\rm BH}}{10^{7}M_{\odot}})= \alpha + \beta \log(\frac{M_*}{10^{10}M_{\odot}}),
  \end {eqnarray}
with  $\alpha = 0.27$, $\beta = 0.98$ based on the local sample of 55 objects measured in a consistent manner. We find that our lensed systems are above the local relation, consistent with the inference from the 32  non-lensed AGNs published by D20 in a  similar redshift range. To quantify the evolution as a function of redshift, we parameterize it as:
\begin{eqnarray}
\label{eq:offset}
\Delta \log \mathcal M_{\rm BH}= \gamma \log (1 + z),
\end{eqnarray} 
where $\Delta \log \mathcal M_{\rm BH}$ is the offset of \mbh\ with respect to the local baseline at fixed \mstar. To make a direct comparison, we reproduce the plot shown in Figure~8 of D20. Then, we add our new lensed AGN measurements and show the result in Figure~\ref{fig:scaling_relation}-(right). We find that the offset from the local relationship is similar for the two samples. Based on Equation~\eqref{eq:offset}, we fit the evolution of the eight lensed systems and obtain $\gamma=1.05\pm0.44$, which agrees with the results of D20 ($\gamma=1.03\pm0.25$ using 32 AGN) within $1-\sigma$ level. The consistency between the two measurements provides an important verification of the accuracy of the results and strengthens the conclusions drawn by D20 that the {\it observed} value of \mbh\ at a fixed \mstar\ tends to be larger at higher redshift than the ones in the local universe. {\bf %It is possible that the uncertainty of the \mstar\ could be higher than our estimation (i.e., 0.2~dex), given that the star formation histories could vary widely at high redshift ($z\sim$2). However, even when our stellar mass uncertainty is boosted, the evolution inferred using the eight lensed systems is still robust --- $\gamma=1.05\pm0.48$ assuming  $\Delta$\mstar = 0.3~dex, and $\gamma=1.03\pm0.53$ assuming $\Delta$\mstar = 0.4~dex.
It is possible that the uncertainty of the \mstar\ could be higher than our assumed 0.2 dex, given that the star formation histories could vary widely at high redshift ($z\sim$2). However, increasing the uncertainty on the stellar mass does not change significantly our conclusions, since the error budget is dominated by the uncertainty in black hole mass. For example, increasing the stellar mass uncertainty to $0.3$ ($0.4$)~dex, we find $\gamma = 1.05\pm0.48$ $(1.03\pm0.53)$.
}

@@

This work is based in part on observations made with the NASA/ESA Hubble Space Telescope, obtained at the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5-26555. X.D., S.B., and T.T. acknowledge support by the Packard Foundation through a Packard Research fellowship to T.T. {\bf This project is received support by the NSF through grant 1907208.} This work is supported by JSPS KAKENHI Grant Number JP18H01251 and the World Premier International Research Center Initiative (WPI), MEXT, Japan. This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No 787886).
C.D.F. acknowledge support for this work from the National Science Foundation under Grant Numbers AST-1312329 and AST-1907396.
SHS thanks the Max Planck Society for support through the Max Planck Research Group.
